\documentclass[12pt, a4paper]{article}
\usepackage{a4wide}
\usepackage{titling}
\usepackage{tabularx}

\setlength{\droptitle}{-10em}
\title{Confidential Computing in Distributed Systems}
\author{Vincent Cui}
\date{\today}

\begin{document}
\maketitle

\section*{Background}

One of the core elements of the digital age is data and information. Often times
more data will lead to a more precise, and thus better usable, information. That
is why it is in the interest of companies to share their data to compute
information. But the problem is that this data is also valuable and could
potentially include sensitive information. So how can a group of mutually
distrusting parties efficiently collaborate in a privacy-preserving manner?

\section*{Assignment}

This thesis shall explore the current state of confidential computing in general
and in the context of distributed computing, discuss different problems that
have to be overcome and investigate various concepts to apply confidential
computing. Another objective will be to implement a proof of concept platform
and define policies that will enable a group of distrusting parties to
collaborate. This platform should be as privacy-preserving as possible while
staying efficient. The thesis will also explain the compromises that will have
to be made in order to balance efficiency and privacy-preservation. It will also
clearly and transparently define trust relationships between the collaborators
and the risks that arise in using the platform.

\section*{Plan}

\begin{tabularx}{\textwidth}{ >{\raggedright\arraybackslash}X
  >{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X }
  \textbf{Phase} & \textbf{Tasks} & \textbf{Time} \\
  \hline\hline
  Research & Collect, sort, and organize literature & 2 weeks \\
  \hline
  Concept & Create a concept that will be implemented & 3 weeks \\
  \hline
  Implementation and Writing & Implement a proof of concept and writing the
  thesis & 6 weeks \\
  \hline
  Correction & Checking literature and sources, proofreading, etc. & 2.5 weeks
  \\
  \hline
  Submission & Printing and submitting the thesis & 0.5 weeks \\
  \hline
  Buffer & Unplanned incidents & 2 weeks \\
  \hline\hline
  \textbf{Sum} & & 16 weeks
\end{tabularx}

\end{document}
