The aim of this thesis was the introduction of a new trusted distributed
computing model that protects the confidentiality of tenant data from untrusted
service providers. Using a threat model that identified threats and issues that
arise when service providers become untrusted, this thesis evaluated the trusted
distributed computing model, which integrates confidential computing and remote
attestation into the traditional distributed computing model. The evaluation
showed that confidential computing combined with remote attestation can solve
the threats and issues of the traditional model.

However, the removal of the service provider as a trusted entity comes with a
trade-off. While the traditional model allowed tenants to shift more and more
responsibilities to the service provider, reducing the complexity of
applications, the trusted model reduces this benefit by giving responsibilities
that could lead to the compromise of the confidentiality of data back to the
application owner.

Additionally, current confidential computing technologies still have performance
limitations that either lessen the cost-efficiency aspect of facilitating
distributed computing systems, or make the use of confidential computing
technology impractical.

Furthermore, confidential computing technologies are still relatively new.
Implementations of TEE platforms, remote attestation protocols, and their
support in other software components are still in early stages of development.
Further research might uncover new vulnerabilities that require changes in
implementations and design choices. Therefore, protocols and implementations are
all still subject to substantial changes. While Cloud providers are eager to
adopt this new technology and are already offering services based on
confidential computing technologies, the two case studies have shown that both
the infrastructure layer and the platform layer are currently still not
fulfilling all requirements of the trusted distributed computing model.

Nevertheless, confidential computing provides very promising hardware-based
primitives that could allow a new generation of trustworthy distributed
computing systems, that meet the increasing demand of more privacy, and enable
projects with stricter security requirements to consider the usage of a shared
distributed computing system.
