Today distributed computing systems are largely based on the Grid or Cloud
computing concept. These systems are based on a layered architecture where each
layer's security is dependent on the security of the layers beneath it, with the
lowest layer being hardware components and the highest layer being applications
that implement abstract business logic. Each layer abstracts the complexity of
the layers below it and provides a simplified interface to the layers above.

Service providers, entities that provide the resources and services of a
distributed computing system, are largely concerned with the protection of the
components that make up the lower layers of distributed computing systems (e.g.
hypervisors providing virtualization of compute resources) from untrusted code
(e.g. virtual machines run by the hypervisor). Traditionally, this isolation
only protects the lower layer from the upper layer, and does not protect the
confidentiality of the upper layer from the lower layer. As a result, the
traditional distributed computing model is completely built around the
assumption that tenants trust the service provider's software stack, including
privileged software such as hypervisors and firmware, but also the service
provider's staff, including system administrators but also those with physical
access to hardware. But this trust is not only present in compute, storage, and
network resources, today many applications utilize services provided by the
service provider that implement privileged functionality, such as authentication
and authorization mechanisms and management of cryptographic key material, in
order to decrease the complexity of the application and be more cost-efficient.

Currently, the best practice for tenants to protect confidential data is to
manage cryptographic keys themselves, for example by operating an own key
management service (KMS) or service provider offered hardware security module
(HSM). While this solution protects cryptographic keys and therefore encrypted
data from the service provider, these keys are still used to decrypt
confidential data that is subsequently used by compute resources managed by the
service provider, making the protection of cryptographic keys from the service
provider pointless. Other solutions rely on new cryptographic primitives that
allow specific compute operations to be directly executed on encrypted data
\cite{monique2013homomorphicencryption, bellare2007searchableencryption}.
However, these solutions currently are either not applicable to general-purpose
computing, or have very high performance limitations that make them impractical.

This thesis introduces the integration of confidential computing technologies
into the traditional distributed computing model. Confidential computing is an
arising technology that provides execution environments, so-called trusted
execution environments (TEEs), that protect applications by utilizing hardware
primitives. By implementing the protection of applications in the lowest layer,
the hardware, it is possible to remove management software such as hypervisors
and operating systems from the list of required trusted software components and
service providers from the list of required trusted entities. While TEEs provide
protection for applications and confidential data, tenants have to be able to
verify that applications are actually running inside TEEs and that the platform
that provides TEEs has not been modified by the service provider. Remote
attestation is an already widely used method for assessing the trustworthiness
of remote devices, components, and environments. Commercially available
confidential computing technologies such as Intel SGX and AMD SEV have built-in
support for producing evidence that can be used for remote attestation.

This thesis will first research the current state of the traditional distributed
computing model, confidential computing technology, and remote attestation
procedures in Chapter \ref{ch:technical-research} and subsequently discuss the
new \textit{trusted distributed computing model} in which service providers are
considered untrusted in Chapter \ref{ch:trusted-distributed-computing-model}.
Section \ref{sec:traditional-distributed-computing} decomposes the two major
distributed computing models Cloud and Grid computing into a generalized
architecture, which is evaluated by Section \ref{sec:untrusted-threat-model} in
order to identify threats from untrusted service providers defining a threat
model. Chapter \ref{ch:trusted-distributed-computing-model} is mainly concerned
with the integration of confidential computing (Section
\ref{sec:confidential-computing}) and remote attestation (Section
\ref{sec:remote-attestation}) into the traditional distributed computing model
and defines requirements of the resulting trusted distributed computing model in
Section \ref{sec:requirements}. Lastly, this thesis will evaluate the untrusted
distributed computing model upon the threat model (Section
\ref{sec:evaluation}), and look at two case studies (Section
\ref{sec:case-studies}).
